---
author: "Megan Foley"
date: "2022-11-26"
output: html_document
---
## qSIP filtering and calcs  
Second script of qSIP calculations used in “Active populations and growth of soil microorganisms are framed by mean annual precipitation in three California annual grasslands”.  

>qSIP functions were downloaded from bitbucket (https://bitbucket.org/QuantitativeSIP/qsip_repo/src/master/) and saved in working directory.  

### Set stuff up  
Load functions for filtering and calculating EAF 18O from qSIP repo
```{r setup, warning=FALSE, message=FALSE}
setwd("~/Documents/winterSIP/wintersip submission/markdowns")
library(kableExtra)
library(tidyverse)

#primary functions
source("qSIP_repo/filter.taxa.R")                     #filter by rep
source("qSIP_repo/all.taxa.calcs.R")                  #compute EAF 18O

#functions called by all.taxa.calcs
source("qSIP_repo/sample.vec.R")                      #sample.vec
source("qSIP_repo/WAD.func.R")                        #WAD.func
source("qSIP_repo/fit.norm.func.R")                   #fit.norm.func
source("qSIP_repo/boot.WAD.func.R")                   #boot.WAD.func
source("qSIP_repo/diff.wad.calc.R")                   #diff.wad.calc
source("qSIP_repo/boot.diff.wad.R")                   #boot.diff.wad
source("qSIP_repo/MW.calc.R")                         #MW.calc
source("qSIP_repo/comparison.message.R")              #comparison.message
source("qSIP_repo/ape.calc.R")                        #ape.calc
source("qSIP_repo/boot.diff.ape.R")                   #boot.diff.ape
source("qSIP_repo/r.calc.R")                          #r.calc
source("qSIP_repo/boot.diff.r.R")                     #boot.diff.r
source("qSIP_repo/boot.TUBE.func.R")                  #boot.TUBE.func
source("qSIP_repo/f.calc.R")                          #f.calc
source("qSIP_repo/boot.diff.f.R")                     #boot.diff.f
```

### Get everything set up & take a look at input files (data.melted & Tcompare)
Load data.melted & Tcompare table. Double check that Soil_Isotope col in data.melted file matches treatment codes in Tcompare file. 
Column classes should be specified as follows:
```{r, echo=FALSE, warning=FALSE, message=FALSE}
col.class.1<- read.csv("input_data/col.class.1.csv")
col.class.2<- read.csv("input_data/col.class.2.csv")
col.class.1 %>% kbl(caption = "data.melted column classes") %>% kable_styling()   
col.class.2 %>% kbl(caption = "Tcompare column classes") %>% kable_styling()   
```
> Note: The Tcompare file tells the qSIP code which tubes to compare to calculate EAF. See Finley et al. 2019 (https://doi.org/10.1007/978-1-4939-9721-3_11) for full description.  
  
> Note: I performed a density correction on the output from the previous script (data.melted file). Script for that will be uploaded soon. The format of data.melted is identical to data.melted.corr.

```{r, message=FALSE,warning=FALSE}
data.melted <- read.csv("~/Documents/winterSIP/wintersip submission/markdowns/input_data/data.melted.corr.csv",
colClasses=c("numeric", "character", "factor", "numeric", "factor", "character", "numeric","numeric", "numeric", "factor", "numeric", "numeric")) %>% select(-X)

Tcompare <- read.table("~/Documents/winterSIP/wintersip submission/markdowns/input_data/trt.comparisons.txt", header=TRUE, sep="\t", colClasses=c("numeric","factor","character","character","character","numeric","character"))

#take a look at the data.melted file
head(data.melted) %>% kbl(caption = "data.melted") %>% kable_styling()

#now make sure Soil_Isotope values match those in Tcompare file
levels(as.factor(data.melted$Soil_Isotope))
Tcompare %>% kbl(caption = "data.melted") %>% kable_styling()      
```

### Filter taxa
Estimates for EAF can be spurious for taxa that occur in very few density fractions and/or reps. We will filter taxa so that only ASVs that occur in at least 2 density fractions (per tube per treatment) and 2 reps (per treatment) are retained.  

* Fraction filtering will be done manually
* Rep filtering willbe done using "filter.taxa" function
```{r, message=FALSE, warning=FALSE}
#Count the number of fractions that an ASV occurs in per tube
data.melted.heavyfilt <- data.melted %>%
  group_by(taxon, Isotope_Treatment, Tube) %>%
  summarise_at(vars(copies.ul), ~sum(. != 0))

#Rename column to "nonzero.count.t.copies.ul" - this signifies how many fractions contain nonzero copies of the taxon & separate data by treatment
names(data.melted.heavyfilt)[4] <- "nonzero.count.t.copies.ul."
head(data.melted.heavyfilt) %>% kbl(caption = "Fractions with nonzero copies per ASV") %>% kable_styling()
join.test <- left_join(data.melted, data.melted.heavyfilt)

#Remove ASVs that occur in less than 2 fractions per tube per treatment
ang<- join.test %>% filter(Soil_Isotope == "Angelo_16" | Soil_Isotope == "Angelo_18") %>% filter(nonzero.count.t.copies.ul. > 1)
hop<- join.test %>% filter(Soil_Isotope == "Hopland_16" | Soil_Isotope == "Hopland_18")%>% filter(nonzero.count.t.copies.ul. > 1)
sed<- join.test %>% filter(Soil_Isotope == "Sedgwick_16" | Soil_Isotope == "Sedgwick_18")%>% filter(nonzero.count.t.copies.ul. > 1)

#Filter so that taxa appear in at least 2 out of 3 biological replicates - I like to do this stepwise to watch how many ASVs are retained
hop.filtered <- filter.taxa(DATA=hop, trt.code.1= Tcompare$trt.code.1[1], trt.code.2= NULL, trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)

hop.filtered.2 <- filter.taxa(DATA=hop.filtered, trt.code.1=NULL, trt.code.2=Tcompare$trt.code.2[1], trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)

ang.filtered <- filter.taxa(DATA=ang, trt.code.1= Tcompare$trt.code.1[2], trt.code.2= NULL, trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)

ang.filtered.2 <- filter.taxa(DATA=ang.filtered, trt.code.1=NULL, trt.code.2=Tcompare$trt.code.2[2], trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)

sed.filtered <- filter.taxa(DATA=sed, trt.code.1= Tcompare$trt.code.1[3], trt.code.2= NULL, trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)

sed.filtered.2 <- filter.taxa(DATA=sed.filtered, trt.code.1=NULL, trt.code.2=Tcompare$trt.code.2[3], trt.refs=NULL, vars=c("taxon", "copies.ul", "Tube", "Soil_Isotope"), min.reps=2)
```

### Run all.taxa.calcs  
This is the function that performs the majority of qSIP calculations and computes estimates of EAF 18O for each ASV by bootstrapping
``````{r, message=FALSE, warning=FALSE, results=FALSE, cache = TRUE}
set.seed(100)
system.time(all.comparisons.hopland <- all.taxa.calcs(X.all=hop.filtered.2, comparisons=Tcompare[1,], vars=c("taxon", "Density_g_ml", "copies.ul", "Tube", "Soil_Isotope", "DNA_ng_ul"), growth.model="exponential", prop.O.from.water=0.6, v.frac=50, copies.cell=6, pgC.cell=0.1, CI=0.95, draws=1000, tailed.test=1))

set.seed(100)
system.time(all.comparisons.angelo <- all.taxa.calcs(X.all=ang.filtered.2, comparisons=Tcompare[2,], vars=c("taxon", "Density_g_ml", "copies.ul", "Tube", "Soil_Isotope", "DNA_ng_ul"), growth.model="exponential", prop.O.from.water=0.6, v.frac=50, copies.cell=6, pgC.cell=0.1, CI=0.95, draws=1000, tailed.test=1))

set.seed(100)
system.time(all.comparisons.sedgwick <- all.taxa.calcs(X.all=sed.filtered.2, comparisons=Tcompare[3,], vars=c("taxon", "Density_g_ml", "copies.ul", "Tube", "Soil_Isotope", "DNA_ng_ul"), growth.model="exponential", prop.O.from.water=0.6, v.frac=50, copies.cell=6, pgC.cell=0.1, CI=0.95, draws=1000, tailed.test=1))
```

### Save output 
```{r, message=FALSE, warning=FALSE, results= 'hold', cache = TRUE}
all.comparisons<- do.call("rbind", list(all.comparisons.hopland, all.comparisons.angelo, all.comparisons.sedgwick))
head(all.comparisons) %>% kbl(caption = "All.comparisons output") %>% kable_styling()
write.csv(all.comparisons, file = "qSIP_output/all.comparisons.csv")
```




